---
title: "Reproducible Research Course Project 1"
author: "Allie Rogers"
date: "November 28, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data

```{r activity}
library("dplyr")
library("magrittr")
activity <- read.csv("activity.csv")
```

## What is the mean total number of steps taken per day?

```{r mean}
activity_daily <- group_by(activity, date) %>%
    summarize(dailysteps = sum(steps))

hist(activity_daily$dailysteps)

mean(activity_daily$dailysteps, na.rm = TRUE)
median(activity_daily$dailysteps, na.rm = TRUE)

```
## What is the average daily activity pattern?
1. Make a time-series plot of the 5-minute interval and the average number 
of steps taken across all days.

```{r daily}
activity_interval <- group_by(activity, interval) %>%
    summarize(avgsteps = mean(steps, na.rm = TRUE))

plot(activity_interval$avgsteps, 
     type = "l",
     main = "Average Daily Activity Pattern",
     ylab = "Average Steps",
     xlab = "Five-Minute Interval")
```

2. Which 5-minute interval contains the maximum number of steps?
```{r max}
activity_max <- filter(activity_interval, 
                       avgsteps == max(activity_interval$avgsteps))
activity_max$interval
```

## Imputing missing values

1. Calculate and report the total number of missing values in the dataset

```{r missing}
nrow(activity) - sum(complete.cases(activity))
```

2. Devise a strategy for filling in all of the missing values in the dataset
3. Create a new dataset that is equal to the original dataset but with the
missing data filled in

```{r filled}
activity_complete <- transform(activity, steps = ifelse(is.na(steps),
                                                    mean(steps, na.rm =  TRUE), 
                                                    steps))
```

4. Make a histogram of the total number of steps taken each day and calculate
and report the mean and median total number of steps per day. 

```{r hist daily}
activity_complete_daily <- group_by(activity_complete, date) %>%
    summarize(dailysteps = sum(steps))
hist(activity_complete_daily$dailysteps)
mean(activity_complete_daily$dailysteps, na.rm = TRUE)
median(activity_complete_daily$dailysteps, na.rm = TRUE)
```

## Are there differences in activity patterns between weekdays and weekends?

1. Create a new factor variable with two levels indicating whether a given
date is a weekend or a weekday day.

```{r factor}
activity_complete <- mutate(activity_complete, 
                            weekday = ifelse(
                           grepl("Saturday|Sunday",weekdays(as.Date(activity_complete$date))),
                                    "weekend", "weekday")) %>%
    mutate(weekday = as.factor(weekday))
```

2. Make a panel plot containing a time series plot of the 5-minute intervals 
and the average number of steps taken, averaged across all weekend or
weekday days

```{r panel plot}
library("ggplot2")

activity_complete_interval <- group_by(activity_complete, interval, weekday) %>%
    summarize(avgsteps = mean(steps, na.rm = TRUE))

g <- ggplot(activity_complete_interval, aes(x = interval, y = avgsteps, color = weekday))
g + geom_line() + facet_grid(. ~ weekday) + ylab("Average Steps")  + 
    xlab("Five-Minute Interval") + ggtitle("Weekday vs. Weekend Activity Pattern")
```